---
title: "ACST 886 Assignment 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1
#### (a)
##### (i) Calculate profit signature 
First step is to get the table of $q_x$, $l_x$ and $_tP_x$.
```{r}
### create a table
year <- c(1,2,3,4)
x <- c()
x[55:58] <- c(55,56,57,58)
qx <- c()
qx[55:57] <- c(0.001046,0.001199,0.001375)
# calculate lx
lx <- c()
lx[55] <- 100000
dx <- c()

for (j in 55:58) {
  dx[j] <- lx[j]*qx[j]
  lx[j+1] <- lx[j]-dx[j]
}

tpx <- c()
for (j in 55:57) {
  tpx[j] <- lx[j+1]/lx[55]
}

table <- cbind(year[1:4],x[55:58],qx[55:58],lx[55:58],tpx[55:58])
colnames(table) <- c("year","x","qx","lx","tpx")
table
```
Next, calculate $a_{56\,:\,\overline{2}|}$ and $a_{57\,:\,\overline{1}|}$, using the formula $a_{x\,:\,\overline{n-1}|}=\frac{v-A_{x\,:\,\overline{n}|}}{d}$
```{r}
### calculate ax_due
A56_3 <- 0.889132
A57_2 <- 0.924607

i <- (1.06/1.0192308)-1
v <- 1/(1+i)
d <- 1-v

ax_due <- c()
ax_due[55:56] <- c((v-A56_3)/d,(v-A57_2)/d)
ax_due[55:56]
```
Now, we need to create vectors for all the cash flows. We have premium, expense at the beginning of year, expense at the end of year, survival annuity. And policy value is calculated by $_{t}V_{x+t\,:\,\overline{n-t}|}=Pa_{x+t\,:\,\overline{n-t}|}$ where P is the annuity payment.
```{r}
## vectors for all cash flows
premium <- c()
premium[55:57] <- c(40250,0,0)

expense_BOY <- c()
expense_BOY[55:57] <- c(100+0.015*40250,0,0)

expense_EOY <- c()
expense_EOY[55:57] <- c(10,10*1.05,10*1.05^2)

annuity_EOY <- c()
annuity_EOY[55:57] <- c(15000,15000*1.0192308,15000*1.0192308^2)

## policy values
policy_EOY <- c()
policy_EOY[55:56] <- annuity_EOY[55:56]*ax_due[55:56]
policy_EOY[57] <- 0

## Create a cashflow table
cashflow <- cbind(year[1:3],x[55:57],premium[55:57],expense_BOY[55:57],
                  expense_EOY[55:57],annuity_EOY[55:57],policy_EOY[55:57])
colnames(cashflow) <- c("year","x","Premium","Expense_BOY",
                        "Expense_EOY","Annuity_EOY","Policy_EOY")
cashflow
```

Now we can calculate the expected cash flows and hence the profit signature.
```{r}
## all the expected cash flows
exp_premium <- c()
exp_premium[55:57] <- premium[55:57]

exp_expense_BOY <- c()
exp_expense_BOY[55:57] <- c(100+0.015*40250,0,0)

exp_expense_EOY <- c()
exp_expense_EOY[55:57] <- tpx[55:57]*c(10,10*1.05,10*1.05^2)

exp_annuity_EOY <- c()
exp_annuity_EOY[55:57] <- tpx[55:57]*c(15000,15000*1.0192308,15000*1.0192308^2)

## expected policy values
exp_policy_EOY <- c()
exp_policy_EOY[55:56] <- tpx[55:56]*policy_EOY[55:56]
exp_policy_EOY[57] <- 0
exp_policy_EOY[54] <- 0

## expected interest 
## earned interest rate is 8%
exp_interest_EOY <- c()
for (j in 55:57) {
  exp_interest_EOY[j] <- (exp_policy_EOY[j-1]+exp_premium[j]-exp_expense_BOY[j])*0.08
}

## profit signature
transfer <- c()
for (j in 55:57) {
  transfer[j] <- exp_policy_EOY[j-1]+exp_premium[j]-exp_expense_BOY[j]-
                 exp_expense_EOY[j]-exp_annuity_EOY[j]+exp_interest_EOY[j]-
                 exp_policy_EOY[j]
}

## a table for all expected cash flows
exp_cashflow <- cbind(year[1:3],x[55:57],exp_premium[55:57],exp_expense_BOY[55:57],
                  exp_expense_EOY[55:57],exp_annuity_EOY[55:57],exp_interest_EOY[55:57],
                  exp_policy_EOY[55:57],transfer[55:57])
colnames(exp_cashflow) <- c("year","x","Exp_Premium","Exp_Expense_BOY",
                        "Exp_Expense_EOY","Exp_Annuity_EOY","Exp_Interest",
                        "Policy_EOY","Transfer")
exp_cashflow
```
The profit signature is the transfer vector.

##### (ii) Measure profitability
Since there is only one initial payment of premium, the NPV of premium is just 40250.
Now calculate the NPV of profit signature.
```{r}
## risk discount rate is 12%
i <- 0.12
EPV_Transfers <- sum(transfer[55:57]/(1+i)^(1:3))
EPV_Transfers
```
Profit margin is calculated by $\frac{EPV(profits)}{EPV(premiums)}$.
```{r}
EPV_Premiums <- sum(exp_premium[55:57]/(1+i)^(0:2))
profit_margin <- EPV_Transfers/EPV_Premiums
profit_margin
```
The profit margin is around 0.5%.

#### (b) IRR
The uniroot function will be used to find the internal rate of return. First we need to create a NPV function in terms of IRR and set NPV to zero to get the required IRR.
```{r}
## create a EPV function in terms of IRR
EPV <- function(irr){
  EPV <- sum(transfer[55:57]/(1+irr)^(1:3))
  return(EPV)
}
## find the IRR 
uniroot(EPV,c(0.4,0.6))
```
The internal rate of return is approximately 50%. We can see that the expected transfer has a large negative value at the first year and followed by two postive values. If we want the NPV of the transfer to be zero, we have to have a large rate to discount back the cash flow.

