---
title: "ACST886 Assignment 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 3
#### (a) Yield on transfers for Q2(a)
```{r}
## Transfer vector from Q2 part (a)
Transfer <- c()
Transfer[30:39] <- c(-38.7603750, 23.778350,21.2621972,18.7990231,16.3686478,
                     13.4784000,10.2381118,6.7380580,3.0520639,-0.7596495)

## create a EPV function in terms of IRR
EPV <- function(irr){
  EPV <- sum(Transfer[30:39]/(1+irr)^(1:10))
  return(EPV)
}
## find the IRR 
uniroot(EPV,c(0.3,0.5))
```
The yield on transfer for Q2 part (a) is about 47.38%.

#### (b) 
##### Double decrement table
First, the independent rate $q_x^d$ and $q_x^w$ are given by Australian insured 2004--8 ultimate (2+) death only non-smoker rates for males.
```{r}
x <- c()
x[30:39] <- c(30:39)
qx.d <- c()
qx.d[30:39] <- c(0.00043,0.00042,0.00041,0.00040,0.00041,
          0.00042,0.00043,0.00045,0.00049,0.00053)
qx.w <- c()
qx.w[30:39] <- c(0.250,0.200,0.175,0.150,0.150,0.150,0.150,0.150,0.150,1.000)
```

The dependent rate is calculated as follow:
$$(aq)_x^d=q_x^d(1-0.5q_x^w)$$
$$(aq)_x^w=q_x^w(1-0.5q_x^d)$$
```{r}
aqx.d <- c()
aqx.d[30:39] <- qx.d[30:39]*(1-0.5*qx.w[30:39])
aqx.d[30:39]
```
```{r}
aqx.w <- c()
aqx.w[30:39]<-qx.w[30:39]*(1-0.5*qx.d[30:39])
aqx.w[30:39]
```

Then calculate $(al)_x$, $(ad)_x^d$ and $(ad)_x^w$.
$$(ad)_x^d=(al)_x*(aq)_x^d$$
$$(ad)_x^w=(al)_x*(aq)_x^w$$
$$(al)_{x+1}=(al)_x-(ad)_x^d-(ad)_x^w$$
```{r}
al.x<-c()
al.x[30]<-100000
ad.x.d<-c()
ad.x.w<-c()

for (j in 30:40) {
  ad.x.d[j]<-al.x[j]*aqx.d[j]
  ad.x.w[j]<-al.x[j]*aqx.w[j]
  al.x[j+1]<-al.x[j]-ad.x.d[j]-ad.x.w[j]
}
al.x[30:40]
```
Combine all those values in a table format.
```{r}
## Double Decrement Table
Decrement <- cbind(x[30:40],qx.d[30:40],qx.w[30:40],aqx.d[30:40],aqx.w[30:40],
                   al.x[30:40],ad.x.d[30:40],ad.x.w[30:40])
colnames(Decrement) <- c("x","qd","qw","(aq)d","(aq)w","(al)","(ad)d","(ad)w")
Decrement
```
##### Possible cash flows for an in-force policy
The vector of premium $P_t$ is
```{r}
## Premium
Pt <- c()
Pt[30:39]<-c(rep(5000,10))
Pt[30:39]
```
The Premium-based fee is $5000(45\%)$ in year 1, $5000(1\%)$ thereafter.
```{r}
## Premium based fee
Prem_base_fee <- c()
Prem_base_fee[30:39] <-c(Pt[30]*0.45,Pt[31:39]*0.01)
Prem_base_fee[30:39]
```
The Buy/Sell margin is $0.5\%P_t(1-45\%)$ in year 1, $0.5\%P_t(1-1\%)$ thereafter.
```{r}
## Buy Sell Margin
BS_Margin <- c()
BS_Margin[30:39] <- c(0.005*Pt[30]*(1-0.45),0.005*Pt[31:39]*(1-0.01))
BS_Margin[30:39]
```
The sum insured is $20,000.
The Insurance Charge is $$Insurance\,charges=1.2q_x^d*SI$$
```{r}
## Sum insured is 20,000
S <- 20000
ins_charge <- c()
ins_charge[30:39] <-1.2*qx.d[30:39]*S
ins_charge[30:39]
```
The Invested premium 
$$IP_t=P_t - Premium\,based\,fees - Buy/Sell\,margins-Insurance\,charges$$
```{r}
IP_in <- c()
IP_in[30:39] <- Pt[30:39]-Prem_base_fee[30:39]-BS_Margin[30:39]-ins_charge[30:39]
IP_in[30:39]
```
The earning rate on the unit fund is $i=10\%$, the asset fee is $m=0.3\%$. 
The value of units at the end of year is $$VU_0=IP_0(1+i)(1-m)\;for\;t=0$$ 
$$VU_t=(VU_{t-1}+IP_t)(1+i)(1-m)\;for\;t=1,2,3,...$$
```{r}
## unit fund earning rate
i <- 0.1
## Asset fees
m <- 0.003
## Value of units
VU_in <- c()
for (j in 31:40) {
  VU_in[30]<-IP_in[30]*(1+i)*(1-m)
  VU_in[j]<-(VU_in[j-1]+IP_in[j])*(1+i)*(1-m)
}
VU_in[30:39]
```
The initial expense rate is 45%, the maintenance expense in year 1 is $75, and the inflation is 25%.
For year 1, the initial expenses at the beginnig of year are $5000(45\%)+75$.
The later expenses at the beginning of year are $75(1+25\%)^{t-1}$ p.a.
```{r}
## Expense
t <-c(1:10)
expense_in <- c()
expense_in[30:39] <- c(Pt[30]*0.45+75,75*1.25^(t[2:10]-1))
expense_in[30:39]
```
Death benefit is $$Death\,Benefits=VU_t+SI$$
```{r}
## Death benefit
Death_benefit_in <- c()
Death_benefit_in[30:39] <- VU_in[30:39]+S
Death_benefit_in[30:39]
```
Surrender benefit is $$Surrender\,Benefits=VU_t$$
```{r}
## Surrender benefit
Surrender_Benefit_in <- c()
Surrender_Benefit_in[30:39] <- VU_in[30:39]
Surrender_Benefit_in[30:39]
```
##### Expected Amounts per Initial Policy for the Unit Fund
Expected invested premium at year t
$$Expected\,Invested\,Premium(EIP_t)=IP_t\frac{(al)_{29+t}}{(al)_{30}}$$
```{r}
## expected invested premium
Exp_IP_in <- c()
Exp_IP_in[30:39] <- IP_in[30:39]*al.x[30:39]/al.x[30]
Exp_IP_in[30:39]
```
Expected value of units at year t
$$Expected\,Value\,of\,Units(EVU_t)=VU_t\frac{(al)_{30+t}}{(al)_{30}}$$
```{r}
## expected value of units
Exp_VU_in <- c()
Exp_VU_in[30:39] <- VU_in[30:39]*al.x[31:40]/al.x[30]
Exp_VU_in[30:39]
```
The the Unit Fund earning rate is 10% p.a.
Expected interest at year t is $i(EVU_{t-1}+EIP_t)$
```{r}
## earning rate is 10#
i <- 0.1
## expected interest rate
Exp_interest <- c()
Exp_interest[30:39] <-c(i*Exp_IP_in[30],i*(Exp_VU_in[30:38]+Exp_IP_in[31:39]))
Exp_interest[30:39]
```
The funds transfered to the Non-unit Fund in year 1 are $$EIP_1+Interst_1-EVU_1$$
The funds transfered to the Non-unit Fund in later years are $$EVU_{t-1}+EIP_t+Interst_t-EVU_t$$
```{r}
Trans_to_Non_unit_in <- c()
for (j in 31:39) {
  Trans_to_Non_unit_in[30]<-Exp_IP_in[30]+Exp_interest[30]-Exp_VU_in[30]
  Trans_to_Non_unit_in[j]<-Exp_VU_in[j-1]+Exp_IP_in[j]+Exp_interest[j]-Exp_VU_in[j]
}
Trans_to_Non_unit_in[30:39]
```
##### Table for Unit Fund
```{r}
year <- c(1:10)

Unit_Fund_in <- cbind(year[1:10],IP_in[30:39],VU_in[30:39],
                  Exp_IP_in[30:39],Exp_interest[30:39],Exp_VU_in[30:39],
                  Trans_to_Non_unit_in[30:39])
colnames(Unit_Fund_in) <- c("year","Invested_Premium","Value_of_Units",
                        "Exp_Invested_Prem","Exp_Interest","Exp_Value_of_Units",
                        "Trans_to_Non_Unit")
Unit_Fund_in
```
##### Expected Amounts per Initial Policy for the Non-unit Fund
Expected premium-based fees at year t
$$Expected\,premium\,based\,fees=premium\,based\,fees\,\frac{(al)_{29+t}}{(al)_{30}}$$
```{r}
## Expected premium-base fee
Exp_prem_based_fee_in <- c()
Exp_prem_based_fee_in[30:39] <- Prem_base_fee[30:39]*al.x[30:39]/al.x[30]
Exp_prem_based_fee_in[30:39]
```
Expected Bul Sell Margin at year t
$$Expected\,Buy\,Sell\,margins=Buy\,Sell\,margins\,\frac{(al)_{29+t}}{(al)_{30}}$$
```{r}
## Expected Buy Sell Margin
Exp_BS_Margin_in <- c()
Exp_BS_Margin_in[30:39] <- BS_Margin[30:39]*al.x[30:39]/al.x[30]
Exp_BS_Margin_in[30:39]
```
Expected insurance charge at year t
$$Expected\,Insurance\,Charges=Insurance\,Charges\,\frac{(al)_{29+t}}{(al)_{30}}$$
```{r}
## expected insurance charge
Exp_ins_charge <- c()
Exp_ins_charge[30:39] <- ins_charge[30:39]*al.x[30:39]/al.x[30]
Exp_ins_charge[30:39]
```
Expected expense at year t
$$Expected\,Expenses=Expenses\,\frac{(al)_{29+t}}{(al)_{30}}$$
```{r}
## expected expense
Exp_expense_in <- c()
Exp_expense_in[30:39]<-expense_in[30:39]*al.x[30:39]/al.x[30]
Exp_expense_in[30:39]
```
Expected death benefit at year t
$$Expected\,Death\,Benefits=Death\,Benefits\,\frac{(ad)_{29+t}^d}{(al)_{30}}$$
```{r}
## expected death benefit
Exp_Death_benefit_in <- c()
Exp_Death_benefit_in[30:39] <- Death_benefit_in[30:39]*ad.x.d[30:39]/al.x[30]
Exp_Death_benefit_in[30:39]
```
Expected surrender benefit at year t
$$Expected\,Surrender\,Benefits=Surrender\,Benefits\,\frac{(ad)_{29+t}^w}{(al)_{30}}$$
```{r}
## expected surrender benefit
Exp_Surrender_Benefit_in <- c()
Exp_Surrender_Benefit_in[30:39]<-Surrender_Benefit_in[30:39]*ad.x.w[30:39]/al.x[30]
Exp_Surrender_Benefit_in[30:39]
```
The Non-unit fund earning rate is 8%.
Expected interest for Non-unit fund at year t
$$Expected\,interest=i(Expected\,premium\,based\,fees+Expected\,Buy\,Sell\,margins+Expected\,Insurance\,Charges-Expected\,Expenses)$$
```{r}
## Non-unit fund earning rate is 8%
i_in_NU <- 0.08
Exp_interst_NU<-c()
Exp_interst_NU[30:39] <- i_in_NU*(Exp_prem_based_fee_in[30:39]+Exp_BS_Margin_in[30:39]+
                          Exp_ins_charge[30:39]-Exp_expense_in[30:39])
Exp_interst_NU[30:39]
```
##### Profit signature
The profit signature, which is he transfers of Non-unit fund are $$Expected\,premium\,based\,fees+Expected\,Buy\,Sell\,margins+Expected\,Insurance\,Charges$$
$$-Expected\,Expenses+Expected\,Interest+Expected\,transfer\,from\,Unit\,Fund$$
$$-Expected\,death\,benefits-Expected\,surrender\,benefits$$
```{r}
## Profit Signature
Transfer_in <- c()
Transfer_in[30:39] <- Exp_prem_based_fee_in[30:39]+Exp_BS_Margin_in[30:39]+
                   Exp_ins_charge[30:39]-Exp_expense_in[30:39]+Exp_interst_NU[30:39]+
                   Trans_to_Non_unit_in[30:39]-Exp_Death_benefit_in[30:39]-
                   Exp_Surrender_Benefit_in[30:39]
Transfer_in[30:39]
```
##### Table for Non-unit Fund
```{r}
Non_Unit_Fund_in <- cbind(year[1:10],Prem_base_fee[30:39],BS_Margin[30:39],
                  ins_charge[30:39],expense_in[30:39],Death_benefit_in[30:39],
                  Surrender_Benefit_in[30:39],Exp_prem_based_fee_in[30:39],
                  Exp_BS_Margin_in[30:39],Exp_ins_charge[30:39],Exp_expense_in[30:39],
                  Exp_interst_NU[30:39],Trans_to_Non_unit_in[30:39],
                  Exp_Death_benefit_in[30:39],Exp_Surrender_Benefit_in[30:39],
                  Transfer_in[30:39])
colnames(Non_Unit_Fund_in) <- c("year","Prem-Based_Fee","BS_Margin","Ins_Charge",
                        "Expense","Death_Bef","Sur_Benefit","Exp_Prem_Based_Fee",
                        "Exp_BS_Margin","Exp_Ins_Charge","Exp_Expense","Exp_Interest",
                        "Trans_from_Unit","Exp_Death_Bef","Exp_Sur_Bef","Transfer")
Non_Unit_Fund_in
```
##### EPV of transfers at 12.5%
$$EPV=\sum_{t=1}^{10}\frac{Transfer_t}{(1+12.5\%)^{t}}$$
```{r}
## EPV of transfer at 12.5%
EPV_Transfer_in <- sum(Transfer_in[30:39]/1.125^t)
EPV_Transfer_in
```
##### Yield on transfers
```{r}
## create a EPV function in terms of IRR
EPV <- function(irr){
  EPV <- sum(Transfer_in[30:39]/(1+irr)^(1:10))
  return(EPV)
}
## find the IRR 
uniroot(EPV,c(-2,1))
```
The new yield on transfers is -192.41%.
